---
title: "Assignment-4"
output: pdf_document
---

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  eval=TRUE,
  echo=FALSE,
  message=FALSE,
  warning=FALSE
)

library(tidyverse)
library(tseries)
library(stargazer)
library(plm)
library(forecast)
theme_set(theme_bw())
```

```{r, code=readLines("Assignment-5-Data.R")}
```

I went with GPD. Here's the data:

```{r}
gdp |> ggplot(aes(x=YEAR, y=GDP)) +
  geom_line()
```

(You can literally *see* COVID-19 on this plot. Same with 2008.)

The "unit of time" in this case is a quarter.

### KPSS Test

$`r kpss$statistic`$, we reject the null (the null being that "the GDP is stationary").

$\neg ``\text{The GDP is stationary.}" \rightarrow ``\text{The GDP is non-stationary.}"$

### Differencing

```{r}
gdp |> ggplot(aes(x=YEAR, y=dGDP)) +
  geom_line()
```

To achieve this result, I had to run differencing twice. (See `Assignment-4-Data.R`.)

### 2${}^\text{nd}$ KPSS Test

$`r kpss2$statistic`$, we do not reject the null.

### Some ARIMA Forecasts

```{r, out.width="50%"}
for (model in models) plot(forecast(model,h=6))
```

### Best ARIMA Forecast (According to AIC)

```{r}
plot(forecast(best.model,h=6))
```